---
title: "a3_task3_anna_talken"
author: "Anna Talken"
date: "10/31/2020"
output: html_document
---

```{r setup, include=FALSE, warnings = FALSE, messages = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
```

### Read in Urchins Data

```{r, warnings = FALSE, messages = FALSE}
sb_urchins <- read_csv(here("data", "urchins.csv")) %>%  # read in urchins data using the here function

clean_names() # clean names using janitor package
```

## Section 1

Here I am exploring the size comparisons between purple and red urchins. 

### Exploratory data visualization 

```{r, warnings = FALSE, messages = FALSE}
urchin_size <- sb_urchins %>% 
  select(size, common_name)
```


### Creating a histogram
```{r, warnings = FALSE, messages = FALSE}
 ggplot(data=urchin_size, aes(x=size)) +
  geom_histogram(bins=10) + 
  facet_wrap(~common_name) +
  labs(x = "Size (cm)", y = "Number")
```

### Creating a boxplot

```{r, warnings = FALSE, messages = FALSE}
ggplot(data=urchin_size, aes(x=common_name, y= size)) +
  geom_boxplot() + 
  labs(x = "Urchin Type", y = "Size (cm)")
```

### Descriptive Statistics Table

```{r, warnings = FALSE, messages = FALSE}
urchin_stats <- sb_urchins %>% 
  group_by(common_name) %>% 
  summarize(mean_size = mean(size),
            median_size = median(size),
            sd_size = sd(size),
            n = n())
urchin_stats
```

### Overall takeaways

Red urchins' average size (mean = 6.7cm) is about 49% greater than purple urchins (mean = 4.5cm). Based on the histograms comparing red and purple urchins, it appears that the red urchins have a normal size distribution whereas the purple urchins are slightly positively skewed. 

## Section 2

Here I am exploring purple urchin counts by site and year.

```{r, warnings = FALSE, messages = FALSE}
purple_urchins <- sb_urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% # Converted date to class "date"
  mutate(year_new = year(date)) %>% # Created new column with just the year 
  mutate(site_full = case_when(site == "NAPL" ~ "Naples Reef", site == "AQUE" ~ "Arroyo Quemado Reef", site == "CARP" ~ "Carpinteria Reef", site == "MOHK" ~ "Mohawk Reef", site == "IVEE" ~ "Isla Vista Reef")) #Creating a new column with the full names of sites
```

### Creating a summary table

This is a summary table with the total number of purple urchins observed by site and year.

```{r, warnings = FALSE, messages = FALSE}
purple_urchin_total <- purple_urchins %>% 
  count(year_new, site_full) 
```

### Making an exploratory graph

```{r, warnings = FALSE, messages = FALSE}
ggplot(data = purple_urchins, aes(x=year_new, fill = site_full)) +
  geom_histogram(bins = 30) + 
  facet_wrap(~site_full) +
    labs(y = "Number of Purple Urchins", x = "Year", fill = "Site")
```

