---
title: "a3_task4_anna_talken"
author: "Anna Talken"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE, messages = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
```

## Read in the Data

```{r, messages = FALSE, warning = FALSE}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names() %>% # clean names
  filter(common_name == "Purple Urchin", site == "NAPL") %>% 
  mutate(date = mdy(date)) %>% # Converted date to class "date"
  mutate(year_new = year(date)) %>% # Created new column with just the year 
  mutate(month_new = month(date)) %>% # Created new column with just the month
  filter(month_new == 1, year_new == 2010) # Filter to only include observations from January in 2010
```

## Exploratory Histogram

```{r, messages = FALSE, warning = FALSE}
ggplot(data=urchins, aes(x=size)) +
  geom_histogram(bins=10) + 
  facet_wrap(~treatment) +
  labs(x = "Size (cm)", y = "Number")
# Creating a histogram to visually compare the size difference between the two treatments (annual vs. control)
```

```{r, messages = FALSE, warning = FALSE}
ggplot(data=urchins, aes(sample=size)) +
  geom_qq() + # Creating a qqplot
  facet_wrap(~treatment) 
```
Based on the histogram, it appears that the two treatment groups don't have a perfectly normal distribution, but are somewhat bell-shaped. The fairly linear trend in the qq plots for both treatments helps to verify that these samples are close to having a normal distribution.  

## Creating a summary table

```{r, messages = FALSE, warning = FALSE}
purple_urchin_stats <- urchins %>% 
  group_by(treatment) %>% 
  summarize(mean_size = mean(size),
            median_size = median(size),
            sd_size = sd(size),
            n = n())

purple_urchin_stats
```
### Split into groups and then pull sample vector 

```{r, messages = FALSE, warning = FALSE}
annual_sample <- urchins %>% 
  filter(treatment == "ANNUAL") %>% 
  pull(size)

control_sample <- urchins %>% 
  filter(treatment == "CONTROL") %>% 
  pull(size)
```

## Run a two sample t-test

```{r, messages = FALSE, warning = FALSE}
t.test(control_sample, annual_sample)
```

After running a two sample t-test, we can see that there is a small difference between the means in each treatment group. The mean of the annual treatment group (mean = 3.916) is only about 5% greater than the mean of the control treatment group (3.725). A p-value of 0.2536 means that there is a 25.36% chance that you could randomly select two samples from populations with the same mean that are this similar by chance.
